---
title: "Lab 07"
author: "Your Name"
date: "10/06/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Comparing base R to tidyverse

So far in this course, you've learned how to manipulate data frames using logical subsetting and other related skills. For this lab, you'll be getting a gentle introduction to data wrangling with dplyr (a tidyverse package).
      -   For odd-numbered questions (through 15), you will use base R (what we've already learned).
      -   For even-numbered questions (through 16), you will use tidyverse functions to accomplish the same task.
      -   This is a little tedious/repetitive! But it's important to understand it works.

For more details on data wrangling with dplyr, see reading <https://r4ds.had.co.nz/transform.html> and RStudio Primer <https://rstudio.cloud/learn/primers/2.2>.

For more details on ggplot2, see <https://r4ds.had.co.nz/data-visualisation.html>.

Please let me know if you spot any errors!

## Assignment

First, load the palmerpenguins package:

```{r}
library(palmerpenguins)
```

1. Use the `head` function (base R) to print the first few rows of the `penguins` dataset.

```{r}

```

2. Use the `glimpse` function (tidyverse) to print the first few rows of the `penguins` dataset.

```{r}

```

3. Extract the rows where the `species` is `Chinstrap` (store these results in an object named `dat`). Then, use base R to print the first few rows of `dat`.

```{r}

```

4. Look at slides 1 through 5 on this flipbook example: <https://evamaerey.github.io/data_manipulation/one_stream_wrangle.html>.
      -   On slide 5, notice how they used `%>%` -- that's just the tidyverse way of passing data into a function.
            -   The `%>%` operator is called a *pipe*, and when you use the pipe operator, it's called *piping*. 
      -   Use the `filter` function to extract the rows of `penguins` where the `species` is `Chinstrap`. Store it in an object named `dat2`.
            -   Hint: your code should look like this: `dat2 <- penguins %>% filter()`. You just have to figure out what to put inside the `filter` function.
      -   On the next line, use tidyverse to print the first few rows of `dat2`.
      
```{r}

```
      
5. In `dat`, `species` is the 1st column. Use base R to remove the `species` column from `dat`. Store the results in `dat`, so that `dat` no longer has a `species` column.
      -   Hint: To remove (instead of subset) a column, use a negative number for the column you are trying to remove.
      -   Then use base R to print the first few rows of `dat`.

```{r}

```

6. Look at slide 6 from the flipbook linked in problem 4. Use the `select` function to remove the `species` column from `dat2`. Then use tidyverse to print the first few rows of `dat2`.

```{r}

```

7. Use base R (`colnames` function) to rename the 4th column of `dat` (`flipper_length_mm`) to `fliplen`. Then use base R to print the first few rows of `dat`.

```{r}

```

8. Look at slides 7 from the flipbook. Use the `rename` function to rename `flipper_length_mm` to `fliplen` (store the output in `dat2`). Then use tidyverse to print the first few rows of `dat2`.

```{r}

```

9. In base R, you can add a column to a data frame using the `cbind` function. For example: `dat <- cbind(dat, new_column = x)`. This will create a new column in `dat` called `new_column`, and the column will contain whatever `x` is. (Alternatively, you could do `cbind(dat,x)` and then use `colnames(dat)[5] <- ` to  rename the new column.
      -   Use the `cbind` function to create a new column in `dat` which contains the product of `bill_length` and `bill_depth`. The column should be named `prod`.
      -   Use base R to print the first few rows of `dat`.
      
```{r}

```

10. Look at slide 8-10 from the flipbook. Use the `mutate` function to create a new column in `dat2` named prod, which contains the product of `bill_length` and `bill_depth`.
      -   Be sure to store the output in `dat2`
      -   Then print the first few rows of `dat2` using tidyverse.

```{r}

```
      
11. Use base R to subset the following columns of `dat`: `bill_length`, `fliplen`, and `prod`. Store the output in `dat`, and use base R to print the first few rows.
      -   Hint: Use the `c()` function to create either a vector of column numbers OR a character vector with the column names (with quotation marks) to subset these columns.

```{r}

```

12. Look at slide 11 from the flipbook. Use the `select` function to overwrite `dat2` so that it only contains the columns `bill_length`, `fliplen`, and `prod`. Then use tidyverse to print the first few rows of `dat2`.
      -   Note: you can use the column names (without quotation marks) OR the column numbers.

```{r}

```

13. In base R, we can use the `sort` function to sort a vector either alphabetically or in numerical order.
      -   For example, `sort(dat$bill_length)` will sort the bill Lengths in the `dat` data frame from lowest to highest.
      -   You can use the `rev` function to reverse a vector. So `rev(sort(dat$bill_length))` would give you these values from highest to lowest.
      -   We can use the `order` function to return the vector positions corresponding to the ordering obtained by sort.
            -   For example, `dat$bill_length[order(dat$bill_length)]` is equivalent to `sort(dat$bill_length)`.
            -   That's not very useful, but it IS useful in reordering the rows of a data frame -- e.g. `dat[order(x),]`
      -   Use the `order` function to reorder the rows of dat based on `bill_length`, **from highest to lowest**. Store the results in `dat` and print the first few rows using base R.
            -   Hint: for highest to lowest ordering, you can use the `rev` function, or you can put a minus sign in front of `dat$bill_length`. Either one will work.

```{r}

```

14. Look at slide 12 from the flipbook. Observe what the `case_when` function does. It's quite interesting! But we're not going to cover it (at least not today). After studying slide 12, move on to slide 13.
      -   Use the `arrange` function to order `dat2` rows by `bill_length` **from highest to lowest**.
      -   Store the output in `dat2` and use tidyverse to print the first few rows.

```{r}

```

15. Using base R, subset `dat` to include rows 3 through 25. Store the results in `dat`, and use base R to print the first few rows.

```{r}

```

16. Look at slide 15 from the flipbook, and use the `slice` function to subset rows 3 through 25. Store the results in `dat2`, and use tidyverse to print the first few rows.

```{r}

```

17. Starting with the penguins dataset, perform all of these changes using a single command. (For stylistic purposes, start a new line AFTER each time you use `%>%`)
      -   At every step, run the code in the console to preview the output to make sure it worked! Then you can add a new `%>%`
      -   Also, don't store this in an object -- just let the entire object print when you run it
      -   Use the `filter` function to extract the rows of `penguins` where the `species` is `Chinstrap`.
      -   Use the `select` function to remove the `species` column
      -   Use the `rename` function to rename `flipper_length_mm` to `fliplen`
      -   Use the `mutate` function to create a new column named prod, which contains the product of `bill_length` and `bill_depth`
      -   Use the `select` function to to only include the columns `bill_length`, `fliplen`, and `prod`.
      -   Use the `arrange` function to order the rows by `bill_length` from highest to lowest
      -   Use the `slice` function to subset rows 3 through 25

```{r}

```

18. Now for some ggplot2 practice! Using the `penguins` dataset and `ggplot2`, make a scatterplot of `bill_length` vs `bill_depth`. Color-code the points by `species`, use point shape to specify `island` and make the size of the points proportional to `flipper_length_mm`.

```{r}

```

19. Make the same plot as you did in 18, but this time:
      -   Don't scale size by `flipper_length_mm`
      -   Use a different them (anything but `theme_grey()`)
      -   Add a line (one for each species) -- within the function that plots the line, include the arguments `method = "lm"` and `se = FALSE`
      -   Make the points transparent (you decide how transparent you want them to be)
      -   Add a title: "Bill Length vs Bill Width"
      -   For the x and y labels, remove the underscore after bill, and also add the units (mm) in parentheses
      
```{r}

```

Save and Knit the document. Be sure the HTML file successfully renders. Then submit BOTH the .Rmd and the HTML file on Webcourses.